A decade of lost precision: Introducing a general framework for image feature matching without geometric constraints

### 2014-04-10 text/journal/paper.tex
State of project by April 10 2014

@@

\title{A general framework for local feature matching in images}

@@

\begin{document}
\maketitle
%
\begin{abstract}
TODO
\end{abstract}
%
\section{Introduction}
%
Matching image points is a crucial ingredient in almost all computer 
vision algorithms that deal with sparse local image features. This 
includes image categorisation \cite{bosch2008scene}, image stitching 
\cite{brown2007automatic}, object detection \cite{zhang2007local} and 
near duplicate detection \cite{zhao2009scale} to mention a few examples.  
These problems all rely on being able to accurately find the 
correspondence(s) of a point on an object in a \emph{query image} given 
one or more \emph{target images} that might contain the same object.  In 
many applications the target images have undergone transformations with 
respect to the query image; in stereo vision, the camera angle is 
different, while in object recognition and near duplicate detection both 
the lighting and even the object itself can also be transformed.

@@

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.85\columnwidth]{images/overview}
\caption{Graphical representation of the \emph{Baseline} and 
\emph{Proposal Set} for different methods in the proposed framework}
\label{fig:overview}
\end{figure}
%
\subsection{A framework of matching methods}
%
\begin{table}[htb]
\caption{Calculating the uniqueness ratio using different sources for 
    the \emph{proposal set} and \emph{baseline set} for the case of two 
images.}
\label{table:distinctness}
	\centering
%	\small
\bgroup
\def\arraystretch{1.5}
    \begin{tabular}{l l | l l}
        %
        %&  \multicolumn{1}{l}{Target}
        %& \multicolumn{1}{l}{Target and Query} \\
\multicolumn{2}{r}{} & \multicolumn{2}{c}{\emph{Proposal Set}} \\
                     & & Target & Target and Query \\
        \cline{2-4}
        \multirow{3}{*}{\rotatebox[origin=c]{90}{\emph{Baseline Set}}} &
  Query            & \emph{Self-Match}   & \emph{Self-Match-Ext} \\
& Target           & \emph{Ratio-Match}   & \emph{Ratio-Match-Ext} \\
& Target and Query & \emph{Mirror-Match} & \emph{Mirror-Match-Ext} \\
        \cline{2-4}
\end{tabular}
\egroup
\end{table}

@@

The three assumptions proposed in order to formalize the order of the 
algorithms have been chosen to reflect conditions under which we usually 
match feature points. However, for each assumption there exists corner 
cases where it might not hold true. In this section we intent to discuss 
these corner cases in order to give an impression of the circumstances 
under which the absolute or relative performance of the algorithms might 
differ.

The first assumption states that there is at most one unique match to 
every feature point which is often the case when in e.g.\ stereo 
matching with natural images. However for recognition of object classes 
or in generated images with repetitive content this assumption is no 
longer valid and using any type of ratio match will lead to incorrect 
results.  As we discussed related to matching more than two images, as 
long as the query image does not contain repetitive objects we can still 
use \emph{Self-Match} without loss of precision even if the target image 
might not contain a unique match.

According to the second assumption, feature descriptors behave such that
when matched, a true correspondence to a query feature will always be 
the nearest neighbor. For actual feature detectors and descriptors 
Mikolajczyk and Tuytelaars shows that this property depends strongly on 
the differences between \emph{query} and \emph{target} image 
\cite{mikolajczyk2005performance} \cite{tuytelaars2008local}. In the 
degenerate case where the closest match is not an actual correspondence, 
it is often the case that other matches are similarly close resulting in 
a uniqueness ratio near 1, effectively filtering out bad 
correspondences.  However when we accept matches with a uniqueness ratio
close to 1, the guarantees about precision and recall equivalence in 
between the different ratio algorithms might no longer hold due to the 
lack of consistent descriptor behaviour. It is worth pointing out that 
if this assumption holds true, matching using only nearest neighbor is 
not necessarily optimal unless there is a complete scene overlap between
the images matched. For cases where there is little or no overlap, a 
nearest neighbor would indiscriminately return false matches even with 
well behaved feature descriptors.

@@

The last assumption states that the uniqueness ratio using either the 
\emph{query} or \emph{target image} as the baseline set would be similar
in distribution when we know that the feature point we are matching has 
a true correspondence. For an ideal feature descriptor where any true 
correspondence between two feature points will have a distance of $0$, 
the third assumption is obviously true for any cases where the second 
closest feature point has a true correspondence in the other image. In 
the general case it is likely that the baseline yields different 
uniqueness ratios depending on the features it contains, but for the 
case where the baseline match has a true correspondence in the other 
image we can verify the assumption by looking at the performance of 
\emph{Self-Match} compared with \emph{Ratio-Match}. To compare we plot 
the actual \emph{uniqueness ratios} measured from 84 3D objects in 
Figure~\ref{fig:ratio_hist}. For ratios lower than $0.7$ this is 
generally the case, but as we approach more lenient thresholds we should
expect the performance of \emph{Self-Match} and \emph{Mirror-Match} to 
decline compared with \emph{Ratio-Match} and \emph{Ratio-Match-Ext}.
% Needs rewording

\section{Experiments and Results}
\label{S:Experiments}
%
\begin{figure}[htb]
    \begin{subfigure}[t]{0.48\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{images/Frog_A}
        \caption{Query Image}
        \label{fig:frog_a}
    \end{subfigure}%
    \quad %add desired spacing between images, e. g. ~, \quad, \qquad a 
    %blank line to force the subfigure onto a new line)
    \begin{subfigure}[t]{0.48\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{images/Frog_B}
        \caption{Auxiliary Image}
        \label{fig:frog_b}
    \end{subfigure}%
    \\ %add desired spacing between images, e. g. ~, \quad, \qquad 
    %a blank line to force the subfigure onto a new line)
    \begin{subfigure}[t]{0.48\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{images/Frog_C}
        \caption{Target Image}
        \label{fig:frog_c}
    \end{subfigure}%
    \quad %add desired spacing between images, e. g. ~, \quad, \qquad 
    %blank line to force the subfigure onto a new line)
    \begin{subfigure}[t]{0.48\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{images/Frog_C_lines}
        \caption{Target Image with epipolar constraints}
        \label{fig:frog_c_lines}
    \end{subfigure}%
    \caption{Creating epipolar constraints based on three source images.  
        In a) the \emph{query image} has been marked with the position 
        of the feature we are attempting to match. In b) the 
        \emph{auxiliary image} is shown. This image is taken with the 
        same rotation as the \emph{query image} but from a higher 
        viewpoint. The line going through the image is the epipolar line 
        of the feature point in the \emph{query image}.  The markers 
        indicate all feature points in the image found near the epipolar 
        line.  In c) we show the \emph{target image} which in this case 
        is rotated 45 degrees from the \emph{query image}.  The line 
        overlaid on the image represents the epipolar line corresponding 
        to the feature point shown in the \emph{query image}.  The 
        markers indicate all feature points in the image found near the 
    epipolar line. In d) we show the \emph{target image} again, this 
time overlaid with the epipolar line corresponding to all the features 
shown in the \emph{auxiliary image}. A true correspondence should be 
found within a small distance of one of the intersections of the line in 
c) and the lines in d). In this particular case both feature points 
shown in c) and d) could possible be a correct correspondence.
}%
    \label{fig:frog}%
\end{figure}%
In this section we will go over the experimental results comparing 
\emph{Self-Match}, \emph{Ratio-Match}, \emph{Ratio-Match-Ext} and 
\emph{Mirror-Match} to each other. We compare the algorithms across 
different image variations and over various descriptors. We use two 
different image databases to compile results over different variations.  
To measure results on non-planar images we test the algorithm on the 3D 
Objects database released by Moreels and Pietro 
\cite{moreels2007evaluation}. The database contains a set of 86 3D 
objects photographed from all sides at 5 degree intervals from two 
different viewpoints and under three different lighting conditions.
To measure the performance over variations such as viewpoint change, 
blur, compression noise, rotation and zoom we also use the graffiti 
image set introduced by Mikolajczyk and Schmid 
\cite{mikolajczyk2005performance}.  This dataset contains 8 scenes each 
with 6 images linked by homographies and for that reason feature only 
planar or almost planar scenes.

The general tests are performed using the \emph{SIFT} descriptor and 
keypoint detector with default parameters as implemented in the OpenCV 
library. To see how the framework performs across descriptors we 
evaluate SURF \cite{bay2006surf}, BRISK \cite{leutenegger2011brisk}, 
BRIEF \cite{calonder2010brief} and FREAK \cite{alahi2012freak} in 
subsection~\ref{label:desc}. Images larger than 1024x768 have been
resized to fit within those dimensions using Imagemagick with default 
parameters and only the luminance channel has been used in the feature 
detection and description stage.

\subsection{Experiments on 3D objects}

@@

The 3D objects dataset allows us to experimentally comparing matching 
algorithms over a large range of object and surface types rotated on a 
turnstile and photographed from every 5 degree turn. We used 84 images 
under three different lighting conditions at 12 different angle 
intervals conducting experiments with a total of 3024 image pairs. To 
validate matches, Moreels and Pietro propose a method to that will 
validate a match as a true or false correspondence \cite[p.  
266]{moreels2007evaluation} using the epipolar constraints.  The method 
is outlined in Figure~\ref{fig:frog}.  According to their experiments 
these constraints are sufficient to reduce the error rate recognizing 
true correspondences to $2\%$. We have used this method
to evaluate the precision and recall of the match algorithms.

To compute the total number of possible correspondences we took each 
feature in a \emph{query image} and counted for how many we could find a 
feature in \emph{target image} which would satisfy the epipolar 
constraints outlined above. In the tests features with no correspondence 
was filtered from the features used to test the matching algorithms to 
avoid matching non-moving background and foreground objects.

We evaluated the 4 matching algorithms on the 3D objects dataset by 
matching images at angular intervals. For each object we picked the 
\emph{query} image as the image taken at 10 degrees rotation for 
calibration stability.  We then proceeded to match this image with the 
same object turned an additional $5, 10, \ldots, 60$ degrees. For every 
angle interval we paired images taken under 3 different lighting 
conditions provided by the dataset.
We included all objects in the database for which angular photos at 5 
degree intervals were available except for the ``Rooster'' and 
``Sponge'' objects due to image irregularities. For each plot we show 
the accumulated results on all 3D objects, weighed by the number of 
possible true correspondences for each. This ensures that each object 
contributes equally to the final result despite some objects
resulting in disproportionally more matches than others.

The results can be seen in Figure~\ref{fig:all_objects}.  
\emph{Ratio-Match} and \emph{Self-Match} show similar results, 
\emph{Self-Match} performing slightly worse for trivial cases and 
improving on \emph{Ratio-Match} for angular differences of more than 30 
degrees. \emph{Mirror-Match} and \emph{Ratio-Match-Ext} clearly 
outperform the others, showing the advantage of composing the set of 
\emph{proposed} features of features from both images. As predicted 
\emph{Mirror-Match} performs better or equal to \emph{Ratio-Match-Ext} 
only rarely showing worse results at high recall. The convergence of all 
algorithms at higher recall is the natural consequence of the 
limitations of the feature descriptors used as noted in the discussion 
of assumptions (Section~\ref{ref:disc_assumptions}). At higher recall we 
start meeting descriptors that are not well behaved meaning that cannot 
rely on the ratio as an indicator of match uniqueness which all ratio 
based algorithms responds equally bad to.

@@

\subsection{Planar Scenes}

To evaluate local feature descriptors, Mikolajczyk and Schmid introduced
a serie of $8$ image sets featuring different transformations such as 
compression noise, rotation and zoom, blur and viewpoint change. All 
images are tied by homographies which makes them convenient for testing 
purposes but as a consequence they all depict planar scenes which is 
somewhat restrictive in terms of image variation. Each scene contains 6 
different images ranged in order of increasing levels of 
transformations. The transformation between image 1 and image 2 in a 
scene is usually only slight while image 6 for the most cases is 
markedly different than image 1.

For each scene we match the first image in the scene with image 3, 4, 5 
and 6 and show the accumulated result weighed by the possible number of 
true correspondences for each pair. We exclude the comparison with image 
2 because it for the most cases is trivial and does not tell us anything 
about the relative performance of the different algorithms. We find 
inliers by using the homography between image pairs as follows: Given a 
potential match between two pixels $p_1$ and $p_2$, $m = \left(p_1, 
p_2\right)$, and a homography $H$ relating the two images $I_1$ and 
$I_2$, we can calculate if $m$ is an inlier by checking if the two 
points satisfy the following criteria:
\begin{equation*}
\left\vert H p_1 - p_2 \right\vert + \left\vert H^{-1}p_2 - p_1 \right\vert < d_{\max}
\end{equation*}
That is, the distance between $p_1$ translated to $I_2$ and $p_2$ 
\emph{plus} the distance between $p_2$ translated to $I_1$ should be 
less than a certain threshold (we use $d_{\max}=10$ pixels here).

Figure~\ref{fig:acr_objects} show the results plotted with precision on 
the y-axis and recall along the x-axis. For scenes such as `bark, `boat' 
(zoom and rotation), `leuven' (light) and `wall' (viewpoint) all 
algorithms show similar results, largely because the precision for all 
algorithms approach 100\% for low recall which does not leave much room 
for improvement. For `bikes' and `trees' (blur) as well as `ubc'
(compression) and `graf' (viewpoint) we see \emph{Mirror-Match} clearly 
outperforming \emph{Ratio-Match}.

@@

We performed the above experiments with the SIFT feature descriptor and 
keypoint detector because this descriptor is the de facto standard in 
local image features. However it is of interest if the improvement gains
shown with the SIFT descriptor translates to other feature descriptors 
such as SURF \cite{bay2006surf} or binary features such as BRIEF 
\cite{calonder2010brief}, BRISK \cite{leutenegger2011brisk} or FREAK 
\cite{alahi2012freak}. To do this we paired each binary descriptor with 
the best performing keypoint detector according to a comparison 
performed by Heinly et al.\ \cite{heinly2012comparative} and the SIFT 
and SURF descriptors with their accompanying keypoint detectors, 
yielding the following matches: BRIEF descriptor with HARRIS keypoint 
detector, BRISK descriptor with FAST keypoint detector. For FREAK we 
used the BRISK feature detector as recommended by Heinly et al. For each 
keypoint detector and feature descriptor pair we used the standard 
implementation from OpenCV 2.4.6.

>>>>

% To compile, make sure to have the following packages installed:
% texlive, texlive-latex-extra, texlive-science, 
% texlive-fonts-recommended, texlive-fonts-extras

@@

\title{A decade of lost precision: Introducing a general framework for 
image feature matching without geometric constraints}

@@

\begin{document}
\maketitle
%
\begin{abstract}
Solutions in computer vision involving the matching of local image 
features frequently use \emph{Ratio-Match} as introduced by Lowe and 
others, but is this really the best approach? We formalize the 
theoretical foundation of \emph{Ratio-Match} and propose a general 
framework encompassing \emph{Ratio-Match} and five other matching 
methods. Using the framework we establish a theoretical order of 
performance in terms of precision and recall between the methods and 
prove that \emph{Ratio-Match} is consistently equalled or outperformed 
by all other methods of the framework.

We confirm the theoretical results experimentally on over 3000 image 
pairs and show that we can easily increase matching precision
without further assumptions about the images we are using. These gains 
comes without increased computation time by changing only a few key 
components of the \emph{Ratio-Match} algorithm.

\end{abstract}
%
\section{Introduction}
%
Matching image points is a crucial ingredient in almost all computer 
vision algorithms that deal with sparse local image features. This 
includes image categorisation \cite{bosch2008scene}, image stitching 
\cite{brown2007automatic}, object detection \cite{zhang2007local} and 
near duplicate detection \cite{zhao2009scale} to mention a few examples.  
These problems all rely on being able to accurately find the 
correspondence(s) of a point on an object in a \emph{query image} given 
one or more \emph{target images} that might contain the same object.  In 
many applications the target images have undergone transformations with 
respect to the query image; in stereo vision, the camera angle is 
different, while in object recognition and near duplicate detection both 
the lighting and even the object itself can also be transformed.

@@

\begin{figure}[b]
\centering
\includegraphics[width=0.85\columnwidth]{images/overview}
\caption{Graphical representation of the \emph{Baseline} and 
\emph{Proposal Set} for different methods in the proposed framework}
\label{fig:overview}
\end{figure}

@@

%
\subsection{A framework of matching methods}
%
\begin{table}[t]
\caption{Calculating the uniqueness ratio using different sources for 
    the \emph{proposal set} and \emph{baseline set} for the case of two 
images.}
\label{table:distinctness}
	\centering
%	\small
\bgroup
\def\arraystretch{1.5}
    \begin{tabular}{l l | l l}
        %
        %&  \multicolumn{1}{l}{Target}
        %& \multicolumn{1}{l}{Target and Query} \\
\multicolumn{2}{r}{} & \multicolumn{2}{c}{\emph{Proposal Set}} \\
                     & & Target & Target and Query \\
        \cline{2-4}
        \multirow{3}{*}{\rotatebox[origin=c]{90}{\emph{Baseline Set}}} &
  Query            & \emph{Self-Match}   & \emph{Self-Match-Ext} \\
& Target           & \emph{Ratio-Match}   & \emph{Ratio-Match-Ext} \\
& Target and Query & \emph{Mirror-Match} & \emph{Mirror-Match-Ext} \\
        \cline{2-4}
\end{tabular}
\egroup
\end{table}

@@

The three assumptions proposed in order to formalize the order of the 
algorithms have been chosen to reflect conditions under which we ideally
would match feature points. However, for each assumption there exists 
corner cases where it is no longer valid. In this section we intent to 
discuss these corner cases in order to give an impression of the 
circumstances under which the absolute or relative performance of the 
algorithms might differ.

The first assumption states that there is at most one unique match to 
every feature point which is often the case when in e.g.\ stereo 
matching with natural images. However for recognition of object classes 
or in generated images with repetitive content this assumption is no 
longer valid and using any assumption of uniqueness will lead to 
incorrect results. However, as long as the query image does not contain 
repetitive objects we can still use \emph{Self-Match} without loss of 
precision even if the target image might not contain a unique match.

According to the second assumption, feature descriptors behave such that
when matched, a true correspondence to a query feature will always be 
the nearest neighbor. For actual feature detectors and descriptors 
Mikolajczyk and Tuytelaars shows that this property depends strongly on 
the differences between \emph{query} and \emph{target} image 
\cite{mikolajczyk2005performance} \cite{tuytelaars2008local}. In the 
degenerate case where the closest match is not an actual correspondence, 
it is often the case that other matches are similarly close resulting in 
a uniqueness ratio near 1, effectively filtering out bad 
correspondences.  This means that with a lenient threshold close to 1, 
the guarantees about precision and recall equivalence in between the 
different ratio algorithms might no longer hold due to the lack of 
consistent descriptor behaviour. It is worth pointing out that if this 
assumption holds true, matching using only nearest neighbor is not 
necessarily optimal unless there is a complete scene overlap between
the images matched. For cases where there is little or no overlap, a 
nearest neighbor would indiscriminately return false matches even with 
well behaved feature descriptors.

@@

The last assumption states that the uniqueness ratio using either the 
\emph{query} or \emph{target image} as the baseline set would be similar
in distribution when we know that the feature point we are matching has 
a true correspondence. For an ideal feature descriptor where any true 
correspondence between two feature points will have a distance of $0$, 
the third assumption is obviously true for any cases where the second 
closest feature point has a true correspondence in the other image. In 
the general case it is likely that the baseline yields different 
uniqueness ratios depending on the features it contains, but for the 
case where the baseline match has a true correspondence in the other 
image we can verify the assumption by looking at the performance of 
\emph{Self-Match} compared with \emph{Ratio-Match}. To compare we plot 
the actual \emph{uniqueness ratios} measured from 84 3D objects in 
Figure~\ref{fig:ratio_hist}. For ratios lower than $0.7$ this is 
generally the case, but as we approach more lenient thresholds the 
performance of \emph{Mirror-Match} might no longer be equal or better 
than \emph{Ratio-Match-Ext} in terms of precision.
% Needs rewording

\section{Experiments and Results}
\label{S:Experiments}
%
\begin{figure}[htb]
    \begin{subfigure}[t]{0.48\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{images/Frog_A}
        \caption{Query Image}
        \label{fig:frog_a}
    \end{subfigure}%
    \quad %add desired spacing between images, e. g. ~, \quad, \qquad a 
    %blank line to force the subfigure onto a new line)
    \begin{subfigure}[t]{0.48\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{images/Frog_B}
        \caption{Auxiliary Image}
        \label{fig:frog_b}
    \end{subfigure}%
    \\ %add desired spacing between images, e. g. ~, \quad, \qquad 
    %a blank line to force the subfigure onto a new line)
    \begin{subfigure}[t]{0.48\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{images/Frog_C}
        \caption{Target Image}
        \label{fig:frog_c}
    \end{subfigure}%
    \quad %add desired spacing between images, e. g. ~, \quad, \qquad 
    %blank line to force the subfigure onto a new line)
    \begin{subfigure}[t]{0.48\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{images/Frog_C_lines}
        \caption{Target Image with epipolar constraints}
        \label{fig:frog_c_lines}
    \end{subfigure}%
    \caption{Creating epipolar constraints based on three source images.  
        In a) the \emph{query image} has been marked with the position 
        of the feature we are attempting to match. In b) the 
        \emph{auxiliary image} is shown. This image is taken with the 
        same rotation as the \emph{query image} but from a higher 
        viewpoint. The line going through the image is the epipolar line 
        of the feature point in the \emph{query image}.  The markers 
        indicate all feature points in the image found near the epipolar 
        line.  In c) we show the \emph{target image} which in this case 
        is rotated 45 degrees from the \emph{query image}.  The line 
        overlaid on the image represents the epipolar line corresponding 
        to the feature point shown in the \emph{query image}.  The 
        markers indicate all feature points in the image found near the 
    epipolar line. In d) we show the \emph{target image} again, this 
    time overlaid with the epipolar lines corresponding to all the 
    features shown in the \emph{auxiliary image}. A true correspondence 
    should be found within a small distance of one of the intersections 
    of the line in c) and the lines in d). In this particular case both 
    feature points shown in c) and d) could possible be a correct 
    correspondence.
}%
    \label{fig:frog}%
\end{figure}%
In this section we will go over the experimental results comparing 
\emph{Self-Match}, \emph{Ratio-Match}, \emph{Ratio-Match-Ext} and 
\emph{Mirror-Match}. We use two different image databases to compile 
results over different variations.  To measure results on realistic and 
variate non-planar images we test the algorithm on the 3D Objects 
database released by Moreels and Pietro \cite{moreels2007evaluation} in 
Section~\ref{S:3dobjects}. The database contains a set of 86 3D objects 
photographed from all sides at 5 degree intervals from two different 
viewpoints and under three different lighting conditions.
To measure the performance over variations such as viewpoint change, 
blur, compression noise, rotation and zoom we also use the graffiti 
image set introduced by Mikolajczyk and Schmid 
\cite{mikolajczyk2005performance} in Section~\ref{S:Planar}.  This 
dataset contains 8 scenes each with 6 images linked by homographies and 
for that reason feature only planar or almost planar scenes.

In the experiments we use the \emph{SIFT} descriptor and keypoint 
detector with default parameters as implemented in the OpenCV library 
v.\ 2.4.6.  To test how the framework performs across descriptors we 
evaluate SURF \cite{bay2006surf}, BRISK \cite{leutenegger2011brisk}, 
BRIEF \cite{calonder2010brief} and FREAK \cite{alahi2012freak} in 
Section~\ref{label:desc}. Across all experiments images larger than 
1024x768 have been resized to fit within those dimensions using 
Imagemagick with default parameters and only the luminance channel has 
been used in the feature detection and description stage.

\subsection{Experiments on 3D objects}
\label{S:3dobjects}

\begin{figure}[t]
    \begin{subfigure}[t]{0.19\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/3d/1}
    \end{subfigure}%
    ~ %
    \begin{subfigure}[t]{0.19\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/3d/2}
    \end{subfigure}%
    ~ %
    \begin{subfigure}[t]{0.19\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/3d/3}
    \end{subfigure}%
    ~ %
    \begin{subfigure}[t]{0.19\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/3d/4}
    \end{subfigure}%
    ~ %
    \begin{subfigure}[t]{0.19\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/3d/5}
    \end{subfigure}%
    \vspace{1.5 mm}

    \begin{subfigure}[t]{0.19\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/3d/6}
    \end{subfigure}%
    ~ %
    \begin{subfigure}[t]{0.19\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/3d/7}
    \end{subfigure}%
    ~ %
    \begin{subfigure}[t]{0.19\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/3d/8}
    \end{subfigure}%
    ~ %
    \begin{subfigure}[t]{0.19\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/3d/9}
    \end{subfigure}%
    ~ %
    \begin{subfigure}[t]{0.19\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/3d/10}
    \end{subfigure}%
    \vspace{1.5 mm}

    \begin{subfigure}[t]{0.19\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/3d/11}
    \end{subfigure}%
    ~ %
    \begin{subfigure}[t]{0.19\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/3d/12}
    \end{subfigure}%
    ~ %
    \begin{subfigure}[t]{0.19\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/3d/13}
    \end{subfigure}%
    ~ %
    \begin{subfigure}[t]{0.19\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/3d/14}
    \end{subfigure}%
    ~ %
    \begin{subfigure}[t]{0.19\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/3d/15}
    \end{subfigure}%
    \vspace{1.5 mm}

    \caption{The first 15 objects from the 3D objects dataset by Moreels
    and Pietro \cite{moreels2007evaluation}}
    \label{fig:3d_objects}
\end{figure}

@@

The 3D objects dataset allows us to experimentally comparing matching 
algorithms over a large range of object and surface types rotated on a 
turnstile and photographed from every 5 degree turn. We used images of 
84 different objects under three different lighting conditions at 12 
different angle intervals conducting experiments with a total of 3024 
image pairs.  To validate matches, Moreels and Pietro propose a method 
to that will validate a match as a true or false correspondence using 
epipolar constraints \cite[p.  266]{moreels2007evaluation}.  Their 
method is outlined in Figure~\ref{fig:frog}.  According to their 
experiments these constraints are sufficient to reduce the error rate 
recognizing true correspondences to $2\%$. We have used their proposed 
method to evaluate the precision and recall of the framework methods.

To compute the total number of possible correspondences we took each 
feature in a \emph{query image} and counted how many of them had a 
feature in the \emph{target image} which would satisfy the epipolar 
constraints outlined above. In the experiments on 3D objects features 
with no correspondences were not included in the set of features used to 
test the matching algorithms to avoid matching non-moving background and 
foreground objects.

We evaluated the 4 matching algorithms on the 3D objects dataset by 
matching images at different angular intervals. For each object we 
picked the \emph{query} image as the image taken at 10 degrees rotation 
for calibration stability.  We then proceeded to match this image with 
the same object turned an additional $d$ degrees, $d \in \{5, 10, 
\ldots, 60\}$.  For every angle interval we compaired images taken under 
3 different lighting conditions as provided by the dataset.
We included all objects in the database for which photos at 5 degree 
angle intervals were available except for the ``Rooster'' and ``Sponge'' 
objects due to image irregularities. For each plot we show the 
accumulated results on all 3D objects, weighed by the number of possible 
true correspondences for the particular object. This ensures that each 
object contributes equally to the final result despite some objects
resulting in disproportionally more matches than others.

The results can be seen in Figure~\ref{fig:all_objects}. In the figure 
we see the performance of the different matching methods in our proposed
framework for $12$ increasingly bigger rotations. The results are shown 
in a precision / recall plot to make it easy to compare performance in 
terms of precision at similar levels of recall.

For objects in the top row rotated between $5$ and $20$ degrees we see a 
gradual increase in the performance gap between the methods as the angle 
increases.  \emph{Ratio-Match} and \emph{Self-Match} show similar 
results while \emph{Mirror-Match} and \emph{Ratio-Match-Ext} outperform 
the them, showing the advantage of composing the \emph{proposed set} of 
features from both images. \emph{Mirror-Match} show performance which is 
consistently slightly better than \emph{Ratio-Match-Ext}. In general we 
see the largest performance improvements at lower recall and a 
convergence of precision at higher recall. This is expected since a 
higher recall is a direct consequence of a more lenient threshold and as
we let the threshold approach $1$ we lose the benefits of thresholding 
on the \emph{uniqueness ratio} the methods all start approximating the 
results of a simple nearest neighbor match. However for all but 
\emph{Ratio-Match} the features with better matches within the same 
image are still weeded out, which explains the tail end of 
\emph{Ratio-Match} at high recall where the other algorithms no longer 
have results. As we demonstrated in \cite{arnfred2013mirror} the removal 
of within-image matches improves performance on image pairs with partial 
or no overlap. For no overlap we would expect a matching algorithm to 
reject matches within regions that have no matching counterparts in the 
other image.

In the middle row featuring rotations of between $25$ and $40$ degrees 
we see the strongest improvement of \emph{Mirror-Match} compared to 
\emph{Ratio-Match} generally peaking at around
a 20 percentage-point increase in precision. Finally the the performance
gain decreases as we further increase the rotation to between $45$ and 
$60$. We suspect that this is partly because assumption \#2 that a 
descriptor is well behaved and match best with a descriptor of a true 
correspondence stops being valid when the images are transformed beyond 
a certain level of recognizability.

To summarize, the methods perform as predicted on the 3D Objects dataset
with \emph{Mirror-Match} in general performing better than 
\emph{Ratio-Match-Ext} which in turn was performing better than 
\emph{Ratio-Match}. The experimental results also show that 
\emph{Self-Match} overall perform equal to or better than 
\emph{Ratio-Match}.

\subsection{Planar Scenes}
\label{S:Planar}

@@

\begin{figure}[t]
    \begin{subfigure}[t]{0.24\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/acr/bark}
    \end{subfigure}%
    ~ %
    \begin{subfigure}[t]{0.24\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/acr/bikes}
    \end{subfigure}%
    ~ %
    \begin{subfigure}[t]{0.24\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/acr/boat}
    \end{subfigure}%
    ~ %
    \begin{subfigure}[t]{0.24\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/acr/graf}
    \end{subfigure}%
    \vspace{1.5 mm}
    % Linebreak
    \begin{subfigure}[t]{0.24\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/acr/leuven}
    \end{subfigure}%
    ~ %
    \begin{subfigure}[t]{0.24\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/acr/trees}
    \end{subfigure}%
    ~ %
    \begin{subfigure}[t]{0.24\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/acr/ubc}
    \end{subfigure}%
    ~ %
    \begin{subfigure}[t]{0.24\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/acr/wall}
    \end{subfigure}%
    \caption{Scenes from the image dataset by Mikolajczyk and Schmid 
    \cite{mikolajczyk2005performance} ordered as they appear in 
Figure~\ref{fig:acr_objects}.}
    \label{fig:acr_images}
\end{figure}

To evaluate local feature descriptors, Mikolajczyk and Schmid introduced
a serie of $8$ image sets featuring different transformations such as 
compression noise, rotation and zoom, blur and viewpoint change. All 
images are tied by homographies which makes them convenient for testing 
purposes but as a consequence they all depict planar scenes which is 
restrictive in terms of image variation. Each scene contains 6 different 
images ranged in order of increasing levels of transformations. The 
transformation between image 1 and image 2 in a scene is only slight 
while image 6 for the most cases is markedly different than image 1.

For each scene we match the first image in the scene with image 3, 4, 5 
and 6 and show the accumulated result weighed by the possible number of 
true correspondences for each pair. We exclude the comparison with image 
2 because it is trivial in most cases and does not tell us much about 
the relative performance of the different methods. We find inliers by 
using the homography between image pairs as follows: Given a potential 
match between two pixels $p_1$ and $p_2$, $m = \left(p_1, p_2\right)$, 
and a homography $H$ relating the two images $I_1$ and $I_2$, we can 
calculate if $m$ is an inlier by checking if the two points satisfy the 
following criteria:
\begin{equation*}
\left\vert H p_1 - p_2 \right\vert + \left\vert H^{-1}p_2 - p_1 \right\vert < d_{\max}
\end{equation*}
That is, the distance between $p_1$ translated to $I_2$ and $p_2$ 
\emph{plus} the distance between $p_2$ translated to $I_1$ should be 
less than a certain threshold (we use $d_{\max}=10$ pixels here).

Figure~\ref{fig:acr_objects} shows the results plotted with precision on 
the y-axis and recall along the x-axis. For scenes such as `bark, `boat' 
(zoom and rotation), `leuven' (light) and `wall' (viewpoint) all methods 
in the framework show similar results, largely because the precision for 
all methods approach 100\% for low recall which does not leave much room 
for improvement. For `bikes' and `trees' (blur) as well as `ubc'
(compression) and `graf' (viewpoint) we see \emph{Mirror-Match} clearly 
outperforming \emph{Ratio-Match} although with a smaller margin than 
with the 3D objects dataset. Because many images in the dataset are 
trivial to match and a few are very difficult we see less spread in 
between the algorithms than with the 3D objects dataset. However the 
theoretical predictions still hold true with \emph{Mirror-Match} 
performing better or equal to \emph{Ratio-Match-Ext} which in turns 
perform better or equal to \emph{Ratio-Match}. As for \emph{Self-Match} 
we are again confirmed that we can achieve performance rivaling 
\emph{Ratio-Match} using only the \emph{query image} as \emph{baseline 
set}.

@@

To evaluate whether the improvement gains shown with the SIFT descriptor 
translates to other feature descriptors such as SURF \cite{bay2006surf} 
or binary features such as BRIEF \cite{calonder2010brief}, BRISK 
\cite{leutenegger2011brisk} or FREAK \cite{alahi2012freak} we evaluated 
the different descriptors on the 3D objects data set.  For the binary 
features they were each paired with the FAST feature detector while SIFT 
and SURF were tested with their respective feature detectors.  For each 
keypoint detector and feature descriptor pair we used the standard 
implementation from OpenCV 2.4.6.

The descriptors were evaluated on 15 images of the 3D object dataset 
with a fixed angle offset of 25 degrees (pictured in 
Figure~\ref{fig:3d_objects}).  The evaluation was otherwise performed as
in the general evaluation of 3D Objects, and the data presented is 
accumulated over the 15 objects, each set of results weighted by the 
possible amount of correspondences.
